---
title: "Exercise 11"
format: html
---

## Load Data

```{r}
library(tidyverse)
library(skimr)
```
```{r}
d <- read_csv("https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/AVONETdataset1.csv", col_names = TRUE)
structure(d)
```
Winnow dataset:
```{r}
d <- d %>% select(Species1, Family1, Order1, Beak.Length_Culmen, Beak.Width, Beak.Depth, Tarsus.Length, Wing.Length, Tail.Length, Mass, Habitat, Migration, Trophic.Level, Trophic.Niche, Min.Latitude, Max.Latitude, Centroid.Latitude, Primary.Lifestyle, Range.Size)
```
```{r}
skim(d)
```
The 7 categorical variables are Species1, Family1, Order1, Habitat, Trophic.Level, Trophic.Niche, and Primary.Lifestyle.

The 12 numeric variables are Beak.Length_Culmen, Beak.Width, Beak.Depth, Tarsus.Length, Wing.Length, Tail.Length, Mass, Migration, Min.Latitude, Max.Latitude, Centroid.Latitude, and Range.Size.

## One Factor ANOVA
#### Step 1
Boxplots of log(Mass) in relation to Trophic.Level and Migration:
```{r}
p1 <- ggplot(data=subset(d, !is.na(Trophic.Level)), aes(x=Trophic.Level, y=log(Mass)))+geom_boxplot(color="blue", fill="blue", alpha=0.2)
p1
```
```{r}
p2 <- ggplot(data=subset(d, !is.na(Migration)), aes(x=as.factor(Migration), y=log(Mass)))+geom_boxplot(color="blue", fill="blue", alpha=0.2)
p2
```
#### Step 2
```{r}
m1 <- lm(log(Mass)~Trophic.Level, d)
summary(m1)
```

```{r}
d$Migration <- as.factor(d$Migration)
m2 <- lm(log(Mass)~Migration, d)
summary(m2)
```

The F-Statistic of each model is large enough to reject the null hypothesis that the F value is equal to zero.

Migration level 1 is the reference level.

Here is the same model with Migration releveled to 3.
```{r}
s <- d
s$Migration<-relevel(s$Migration, ref=3)
m3 <- lm(log(Mass)~Migration, s)
summary(m3)
```

#### Step 3

```{r}
m <- aov(log(Mass) ~ Migration, data = d)
posthoc <- TukeyHSD(m, which = "Migration", ordered = TRUE, conf.level = 0.95)
plot(posthoc)
```
The 2-1 categories differ significantly from each other.

#### Step 4



