---
title: "Exercise 11"
format: html
---

## Load Data

```{r}
#| warning: false
library(tidyverse)
library(skimr)
library(infer)
library(broom)
```

```{r}
#| message: false
d <- read_csv("https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/AVONETdataset1.csv", col_names = TRUE)
```
Winnow dataset:
```{r}
d <- d %>% select(Species1, Family1, Order1, Beak.Length_Culmen, Beak.Width, Beak.Depth, Tarsus.Length, Wing.Length, Tail.Length, Mass, Habitat, Migration, Trophic.Level, Trophic.Niche, Min.Latitude, Max.Latitude, Centroid.Latitude, Primary.Lifestyle, Range.Size)
skim(d)
```

The 7 categorical variables are Species1, Family1, Order1, Habitat, Trophic.Level, Trophic.Niche, and Primary.Lifestyle.

The 12 numeric variables are Beak.Length_Culmen, Beak.Width, Beak.Depth, Tarsus.Length, Wing.Length, Tail.Length, Mass, Migration, Min.Latitude, Max.Latitude, Centroid.Latitude, and Range.Size.

## One Factor ANOVA
#### Step 1
Boxplots of log(Mass) in relation to Trophic.Level and Migration:
```{r}
p1 <- ggplot(data=subset(d, !is.na(Trophic.Level)), aes(x=Trophic.Level, y=log(Mass)))+geom_boxplot(color="blue", fill="blue", alpha=0.2)
p1
```
```{r}
p2 <- ggplot(data=subset(d, !is.na(Migration)), aes(x=as.factor(Migration), y=log(Mass)))+geom_boxplot(color="blue", fill="blue", alpha=0.2)
p2
```
#### Step 2
```{r}
m1 <- lm(log(Mass)~Trophic.Level, d)
summary(m1)
```
While the F Stat shows that the model is significant, the "Omnivore" trophic level does not significantly relate to log(Mass).

```{r}
d$Migration <- as.factor(d$Migration)
m2 <- lm(log(Mass)~Migration, d)
summary(m2)
```

The F-Statistic of each model is large enough to reject the null hypothesis that the F value is equal to zero.

Migration level 1 is the reference level.

Here is the same model with Migration releveled to 3.
```{r}
s <- d
s$Migration<-relevel(s$Migration, ref=3)
m3 <- lm(log(Mass)~Migration, s)
summary(m3)
```

#### Step 3

```{r}
m <- aov(log(Mass) ~ Migration, data = d)
posthoc <- TukeyHSD(m, which = "Migration", ordered = TRUE, conf.level = 0.95)
plot(posthoc, xlim=c(-0.5, 1))
```
All levels of migration are significantly different from the null, which would hold if the confidence intervals intersected with zero. The 2-1 categories differ the most from each other compared to other category combinations.

#### Step 4
create log.Mass variable because the *specify* function does not like log(Mass)
```{r}
d <- d %>%
  mutate(log.Mass=log(Mass))
```
Permute the F stat with the {infer} package
```{r}
permuted.F <- d %>%
  specify(log.Mass ~ Trophic.Level) %>%  # specify model
  hypothesize(null = "independence") %>% # null hypothesis of independence
  generate(reps = 1000, type = "permute") %>%  # generate permutations
  calculate(stat = "F") # calculate the F statistic for the AOV
```
Create an "original" F stat using aov function
```{r}
original.F <- aov(data = d, log.Mass ~ Trophic.Level) %>%
  tidy() %>%
  filter(term=="Trophic.Level")
visualize(permuted.F) +
  shade_p_value(obs_stat=original.F$statistic, direction="greater")
```
Generate p value
```{r}
p.value <- permuted.F %>%
  get_p_value(obs_stat = original.F$statistic, direction="greater")
p.value
```

The p value is zero, allowing us to reject the null hypothesis that there is no difference in log(Mass) between the trophic level groups.

## Challenge 2

#### Step 1

Get residuals, add to dataframe
```{r}
rbl <- lm(log(Beak.Length_Culmen)~log(Mass), d)
rtl <- lm(log(Tarsus.Length)~log(Mass), d)

d <- d %>% mutate(
  rel.beak.length=rbl$residuals,
  rel.tarsus.length=rtl$residuals
)
head(d)
```
#### Step 2
Relative Tarsus Length vs. Primary Lifestyle
```{r}
p3 <- ggplot(data=subset(d, !is.na(Primary.Lifestyle)), aes(x=Primary.Lifestyle, y=rel.tarsus.length))+geom_boxplot(color="blue", fill="blue", alpha=0.2)
p3
```
Relative Beak Length vs. Trophic Niche
```{r}
p4 <- ggplot(data=subset(d, !is.na(Trophic.Niche)), aes(x=Trophic.Niche, y=rel.beak.length))+geom_boxplot(color="blue", fill="blue", alpha=0.2)
p4
```

#### Step 3

```{r}
m.aov <- aov(data = subset(d, !is.na(Migration)), log(Range.Size) ~ Migration)
summary(m.aov)
```
The range size does appear to be significantly related to the migration categorization. 
```{r}
m.lm <- lm(log(Range.Size) ~ Migration, data = subset(d, !is.na(Migration)))
tidy(m.lm)
```
Categories 2 and 3 are significantly different from reference level of 1. 
Relevel to 3:
```{r}
d$Migration<-relevel(d$Migration, ref=3)
m.lm2 <- lm(log(Range.Size) ~ Migration, data = subset(d, !is.na(Migration)))
tidy(m.lm2)
```
Categories 1 and 2 are significantly different from the reference level of 3. 

Post-Hoc Tukey HSD:
```{r}
posthoc2 <- TukeyHSD(m.aov, which = "Migration", ordered = TRUE, conf.level = 0.95)
plot(posthoc2, xlim=c(-1,3))
```

The post-hoc Tukey Honest Significant Differences test shows the largest difference between categories is between 1 and 3. All categories, however, are significantly different from one another.

#### Step 4
Get Passeriformes:
```{r}
pass <- d %>% filter(Order1=="Passeriformes")
```

##### Relative Beak Length vs Primary Lifestyle

```{r}
p5 <- ggplot(data=subset(pass, !is.na(Primary.Lifestyle)), aes(x=Primary.Lifestyle, y=rel.beak.length))+geom_boxplot(color="blue", fill="blue", alpha=0.2)
p5
```

```{r}
passRBL_PL <- aov(data = subset(pass, !is.na(Primary.Lifestyle)), rel.beak.length ~ Primary.Lifestyle)
summary(passRBL_PL)
```

```{r}
passRBL_PL.lm <- lm(rel.beak.length ~ Primary.Lifestyle, data = subset(pass, !is.na(Primary.Lifestyle)))
summary(passRBL_PL.lm)
```
##### Relative Beak Length vs Trophic Level

```{r}
p6 <- ggplot(data=subset(pass, !is.na(Trophic.Level)), aes(x=Trophic.Level, y=rel.beak.length))+geom_boxplot(color="blue", fill="blue", alpha=0.2)
p6
```

```{r}
passRBL_TL <- aov(data = subset(pass, !is.na(Trophic.Level)), rel.beak.length ~ Trophic.Level)
summary(passRBL_TL)
```

```{r}
passRBL_TL.lm <- lm(rel.beak.length ~ Trophic.Level, data = subset(pass, !is.na(Trophic.Level)))
summary(passRBL_TL.lm)
```

#### Step 5

Check for similar variances:
```{r}
stats <- pass %>% group_by(Primary.Lifestyle, Trophic.Level) %>%
  summarize("mean(rel.beak.length)" = mean(rel.beak.length),
    "sd(rel.beak.length)"= sd(rel.beak.length))

max(stats$`sd(rel.beak.length)`)/min(stats$`sd(rel.beak.length)`)
```
```{r}
# check that variances in each group are roughly equal (ratio of max/min is <2)
p7 <- ggplot(data=subset(pass, !is.na(Trophic.Level)), aes(y = rel.beak.length, x = Primary.Lifestyle)) + geom_boxplot() +
  facet_wrap(~Trophic.Level, ncol=4) +
  xlab("Primary Lifestyle") + ylab("Relative Beak Length")
# and let's plot what the data look like
# p <- p + geom_point() # uncommenting this shows all points
#p <- p + stat_summary(data=d, aes(y = femur_length, x = sex),
 # fun = base::mean,
  # make sure we use {base} version of mean
 # color = "darkgreen", geom = "point", shape = 8, size = 6)
#p
p7
```

Check each ANOVA separately:
```{r}
summary(aov(data=pass, rel.beak.length ~ Primary.Lifestyle))
```
```{r}
summary(aov(data=pass, rel.beak.length ~ Trophic.Level))
```
There seems to be a significant relationship for both variables.

##### Multiple ANOVA:

```{r}
pass_aov_mult <- aov(data = pass, rel.beak.length ~ Primary.Lifestyle+Trophic.Level)
summary(pass_aov_mult)
```
This model result suggests that both variables are significant predictors of relative beak length. The trophic level of each species appears to have a larger impact on this variable than does primary lifestyle, although the difference is not very large.

#### Step 6

```{r}
pass_aov_int <- aov(data = pass, rel.beak.length ~ Primary.Lifestyle+Trophic.Level+Primary.Lifestyle:Trophic.Level)
summary(pass_aov_int)
```

After considering interactions, it seems that while the two independent categories are still the best predictors of relative beak length, a small but significant interaction between the two categories also contributes to the result. 

#### Step 7

```{r}
interaction.plot(
  x.factor = pass$Primary.Lifestyle,
  xlab = "Primary Lifestyle",
  trace.factor = pass$Trophic.Level,
  trace.label = "Trophic Level",
  response = pass$rel.beak.length,
  fun = base::mean, # make sure we use {base} version
  ylab = "Relative Beak Length"
)
```
Where the interaction plot is parallel, there is no interaction. This is seen well in comparing the **omnivore** and **herbivore** plots. However, the **carnivore** plot shows interaction with the **omnivore** plot, as they are not parallel and intersect. 
